---
title: "Lab 06 - Ugly charts and Simpson's paradox"
author: "Ryan Wheat"
date: "02/23/23"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
library(mosaicData) 
```

```{r load-data}
staff <- read_csv("data/instructional-staff.csv")
```


### Exercise 1

```{r wide-to-long}

staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))

#plot long dataframe

staff_long %>%
  ggplot(aes(x = year,
             y = value,
             group = faculty_type,
             color = faculty_type)) +
  geom_line() +
  labs(title = "Instructional Staff Employment Plans", x = "Year", y = "Percentage of Hires", color = "Faculty Type")

```


### Exercise 2

This graph works, but given the question that we need to answer, I would collapse the full-time faculty into one category so that there is less clutter -- and the visual can be assessed in terms of full time faculty vs. part time faculty.


### Exercise 3

These visuals are truly horrendous: I cannot figure out what the researchers are trying to convey. These are the changes I'd make: 

- Label which color on the plot represents fish captured vs. fish farmed in each country. 

- I'm not sure it makes sense to put both of these variables on the same graph, becuase the amount of fish farmed in China is so large, it makes it hard to tell anything else. So scaling things more appropriately for each variable would be useful.

- On that note, it is really interesting to me that China alone farms more fish than the rest of the world combined. To highlight this, I'd like to make some sort of visual that compares China's fish farming stats to that of every other country.

```{r fisheries-visual}

fisheries <- read_csv("data/fisheries.csv")

#label for captured vs. farmed

fisheries %>%
  ggplot(aes(x = country, y = capture)) +
  geom_density() +
  labs(title = "Total Fish Captured per Country", x = "Country", y = "Total Fish Captured")

```


### Exercise 4

```{r Wickham-study}

data("Whickham")

?Whickham
library(performance)
performance::compare_performance()
```



...

Add exercise headings as needed.
