---
title: "Lab 06 - Ugly charts and Simpson's paradox"
author: "Ryan Wheat"
date: "02/23/23"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
library(mosaicData) 
```

```{r load-data}
staff <- read_csv("data/instructional-staff.csv")
```


### Exercise 1

```{r wide-to-long}

staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))

#plot long dataframe

staff_long %>%
  ggplot(aes(x = year,
             y = value,
             group = faculty_type,
             color = faculty_type)) +
  geom_line() +
  labs(title = "Instructional Staff Employment Plans", x = "Year", y = "Percentage of Hires", color = "Faculty Type")

```


### Exercise 2

This graph works, but given the question that we need to answer, I would collapse the full-time faculty into one category so that there is less clutter -- and the visual can be assessed in terms of full time faculty vs. part time faculty.


### Exercise 3

These visuals are truly horrendous: I cannot figure out what the researchers are trying to convey. These are the changes I'd make: 

- Label which color on the plot represents fish captured vs. fish farmed in each country. 

- I'm not sure it makes sense to put both of these variables on the same graph, becuase the amount of fish farmed in China is so large, it makes it hard to tell anything else. So scaling things more appropriately for each variable would be useful.

- On that note, it is really interesting to me that China alone farms more fish than the rest of the world combined. To highlight this, I'd like to make some sort of visual that compares China's fish farming stats to that of every other country.

```{r fisheries-visual}

fisheries <- read_csv("data/fisheries.csv")

#label for captured vs. farmed

fisheries %>%
  ggplot(aes(x = country, y = capture)) +
  geom_density() +
  labs(title = "Total Fish Captured per Country", x = "Country", y = "Total Fish Captured")

```


### Wickham

#### Exercise 1 

This seems like observational data, because it seems extremely unethical to manipulate ppl to be smokers experimentally. 

#### Exercise 2

There are 1314 observations. Each observation represents one participant.

#### Exercise 3

There are 3 variables: outcome, smoking status, and age.

```{r Wickham-study-visual}

data("Whickham")

?Whickham
library(performance)

#performance::compare_performance()

#plot each variable

Whickham %>%
  ggplot(aes(x = outcome)) +
  geom_bar() + theme_bw() + labs(x = "Currently Dead or Alive?", y = "Count")

Whickham %>%
  ggplot(aes(x = smoker)) +
  geom_bar() + theme_bw() + labs(x = "Smoker at Baseline?", y = "Count") 

Whickham %>%
  ggplot(aes(x = age)) +
  geom_histogram() + theme_bw() + labs(x = "Age at Baseline", y = "Count")

```

#### Exercise 4

My very novel and revolutionary prediction is that smokers at baseline will have worse health/mortality outcomes 20 years later.

#### Exercise 5

76% of smoking individuals (at the time of the intake survey) were still alive 20 years later. 69% of non-smoking individuals were still alive after 20 years. This pattern of results seems to suggest that smoking does not decrease life expectancy -- which we know can't be right.

```{r Wickham-smoker-outcome}

#visualization

Whickham %>%
  ggplot(aes(x = smoker, fill = outcome)) + geom_bar() + theme_bw() + labs(x = "Smoking at Time of Intake", y = "Count", fill = "Mortality")

#conditional probability

Whickham %>%
  count(smoker, outcome) %>%
  group_by(smoker) %>%
  mutate(mortal_prob = n/sum(n))

```

#### Exercise 6

```{r age-category}

#create age category variable

Whickham <- Whickham %>%
  mutate(age_cat = case_when(age <= 44 ~ "18-44",
age > 44 & age <= 64 ~ "45-64",
age > 64 ~ "65+"))

```

#### Exercise 7

```{r better-Wickham-visual}

#visualization

Whickham %>%
  ggplot(aes(x = smoker, fill = outcome)) + geom_bar() + facet_wrap(~ age_cat) + theme_bw() + labs(x = "Smoking at Time of Intake", y = "Count", fill = "Mortality")

#conditional probability

Whickham %>%
  count(smoker, age_cat, outcome) %>%
  group_by(smoker) %>%
  mutate(mortal_prob = n/sum(n))

```



...

Add exercise headings as needed.
